---
title: "Precompiled vignettes and articles with rawvignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Precompiled vignettes and articles with rawvignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)

# This vignette demonstrates the workflow inside a disposable throwaway
# package created in a temporary directory. We create it here and point
# knitr's root directory at it, so every chunk below runs as if from that
# package's root. (opts_knit$set(root.dir = ...) is the supported way to do
# this; a setwd() inside a chunk is undone by knitr after that chunk and
# would not persist to later chunks.)
demo_pkg <- file.path(tempdir(), "demopkg")
unlink(demo_pkg, recursive = TRUE)
dir.create(demo_pkg, showWarnings = FALSE)
writeLines(c("Package: demopkg", "Version: 0.0.1"),
           file.path(demo_pkg, "DESCRIPTION"))
knitr::opts_knit$set(root.dir = demo_pkg)
```

```{r}
library(rawvignette)
```


By default, R rebuilds your vignettes every time the package is built --
during `R CMD check`, on CRAN, on GitHub Actions, and when pkgdown
generates your site. Usually that is a good thing: it keeps the rendered
output honest. 

## The problem

But when a vignette (or a pkgdown article) does heavy work -- a long or 
memory-intensive computation, a call to a rate-limited API, a large 
dataset -- rebuilding it everywhere, every time, can become annoying or 
impossible.

Here is the kind of chunk that causes the trouble. Imagine the `Sys.sleep()` 
below is an intensive simulation that depends on a large input dataset and 
takes two hours to run:

```{r slow-example}
expensive_result <- {
  Sys.sleep(2)  # stand-in for genuinely slow work
  summary(1:1000)
}
expensive_result
```

If that chunk lives in an ordinary vignette, every build server needs access
to the input data and has to run a two-hour job each time the vignette is built. 


## The idea

The goal of `rawvignette` is to build the vignette once, on your machine, and ship 
the already-computed result.

The workflow splits a vignette into two stages:

- A **source** in `vignettes-raw/`, with live, executable code. You edit
  and run this locally, whenever you choose.
- A **staged** `.Rmd` in `vignettes/`, produced by knitting the source.
  Its code has already been executed and its output inlined, so there is
  nothing expensive left to run.

Downstream tools only ever see the staged form. `R CMD check`, CRAN, and
pkgdown re-knit it, but since its chunks are already evaluated, that
re-knit is instant.

This is the well-established "precompiled vignette" idea (see the
[rOpenSci tech note](https://ropensci.org/blog/2019/12/08/precompute-vignettes/));
`rawvignette` just adds tooling around it -- scaffolding, figure handling,
`.Rbuildignore` management, and a freshness check -- so you do not have to
script those by hand.


## Walking through the workflow

The functions expect to run from a package root. To keep this vignette
self-contained and runnable, the examples below operate in a disposable
throwaway package created in a temporary directory (set up invisibly when
this vignette starts). In your own work you would skip that step and simply
run the workflow from your package's top level.

### 1. Scaffold a source

`use_raw_vignette()` sets up a skeleton for a new pre-compiled vignette or article.
You can also use it to migrate an existing standard vignette or article to the
`rawvignette` workflow. The function wires up the supporting infrastructure 
(the `vignettes-raw/` directory, a `.Rbuildignore` entry, and a `vignettes/figures/` 
directory for vignettes), and either creates a new vignette/article `.Rmd` file for you
to edit, or moves your existing `.Rmd` into the new directory.

```{r scaffold}
use_raw_vignette("intro", title = "Introduction")
```

That leaves you with a skeleton source to edit:

```{r show-source}
list.files("vignettes-raw")
```

In real use you would now open `vignettes-raw/intro.Rmd` and write your
vignette there, iterating and running code the same way you would for a 
normal vignette. For the demo, let's drop in a chunk that does some 
(pretend-expensive) work:

```{r write-source}
writeLines(c(
  "---",
  'title: "Introduction"',
  "output: rmarkdown::html_vignette",
  "vignette: >",
  "  %\\VignetteIndexEntry{Introduction}",
  "  %\\VignetteEngine{knitr::rmarkdown}",
  "  %\\VignetteEncoding{UTF-8}",
  "---",
  "",
  "```{r}",
  "# Pretend this is slow:",
  "nrow(mtcars)",
  "```"
), "vignettes-raw/intro.Rmd")
```


### 2. Precompile

`precompile_raw_vignettes()` knits every source under `vignettes-raw/`
to its counterpart in `vignettes/`, running the code once and inlining
the results. It also injects a "do not edit by hand" notice identifying
the source.

```{r precompile}
precompile_raw_vignettes()
```

The shipped vignette now exists, with the computed output baked in:

```{r show-output}
list.files("vignettes")
out <- readLines("vignettes/intro.Rmd")
cat(head(out, 20), sep = "\n")
```

Notice the generated notice at the top: editing `vignettes/intro.Rmd`
directly would be a mistake, because the next precompile overwrites it.
The source of truth is `vignettes-raw/intro.Rmd`.

### 3. Check freshness

Because the two files can drift -- you edit the source, then forget to
re-precompile -- you can use `check_raw_vignettes()` to compares their 
file modification times and flags any stale output.

```{r check-fresh}
check_raw_vignettes()
```

Right after precompiling, everything is fresh. If you now edit the source
without re-precompiling, the check notices:

```{r check-stale}
# Simulate editing the source after the output was built.
Sys.setFileTime("vignettes-raw/intro.Rmd", Sys.time() + 5)
check_raw_vignettes()
```

This is a heuristic, not a guarantee. It catches the common slip -- an
un-precompiled edit -- but cannot detect staleness caused by changes to
your package code, data files, or dependency versions, since those do not
touch the source `.Rmd`. The only robust guarantee is to re-run
`precompile_raw_vignettes()` before a release.

To catch the forgot-to-precompile slip automatically, you can install a
git pre-commit hook that runs the freshness check and blocks a commit when
an output is stale, using `use_raw_vignette_hook()`. If you already have a 
pre-commit hook (for example, the one
`usethis::use_readme_rmd()` installs), pass `on_collision = "append"` to
add the check alongside it instead of replacing it. The hook is a local
convenience: it lives in `.git/hooks/`, so it is not version-controlled and
must be re-installed per clone, and like the freshness check it compares
modification times only. See `?use_raw_vignette_hook` for details.


## Articles

pkgdown also displays **articles**: vignette-like documents that appear on
your website but do not ship with the package (handy for content that is
large, depends on unavailable resources, or simply does not belong on
CRAN). `rawvignette` precompiles articles the same way -- the main
difference is where they live.

Pass an `articles/` prefix, and the source goes to
`vignettes-raw/articles/`, the output to `vignettes/articles/`, and an
extra `.Rbuildignore` entry keeps the whole `vignettes/articles/` subtree
(rendered article plus its figures) out of the package tarball:

```r
use_raw_vignette("articles/benchmark", title = "Benchmark")
```

This creates an `Rmd.` with slightly different header parameters, appropriate
for a pkgdown article. Everything else -- editing, precompiling, the freshness 
check -- works identically. There is no separate flag to set; the `articles/` 
location is the single signal that a document is an article rather than a vignette.


## When to use `rawvignette`

Precompilation has a real cost: your documentation no longer rebuilds
automatically when the package changes, so you take on the discipline of
re-precompiling when relevant inputs change. That trade is worth it for
genuinely expensive vignettes, and not worth it for cheap ones. Precompile
the documents that are slow, flaky, or resource-hungry; leave the rest as
ordinary vignettes.
